---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

here text should follow

## What is mean total number of steps taken per day?

and here

## What is the average daily activity pattern?

and here

## Imputing missing values

and here

## Are there differences in activity patterns between weekdays and weekends?

and here

---> all from here needs to be replaces!!!!!

```{r libraries}
library(lubridate)
library(stringr)
library(dplyr)
library(lattice)
```
Load data and store it in data.frame *df*

```{r}
df <- read.csv("/home/remco/Desktop/data/activity.csv")

str(df)
```

Now change the *interval* column from numeric to character.  
After that leading zeros are added so date-time stamps can be calculated.

```{r}
df$interval2 <- str_pad(df$interval,4,pad="0")
str(df)
```

Now using the *date* column and the reformatted *interval* column, a datetime value can be calculated. For this we use **ymd_hm** from the **lubridate** library.

```{r}
library(lubridate)
df$datetime <- ymd_hm(paste(df$date,df$interval2, sep=" "))

str(df)
```

```{r}
days <- group_by(df, date)
steps_per_day <-summarize( days, steps = sum(steps, na.rm=TRUE))
hist(steps_per_day$steps, breaks=30)

mean(steps_per_day$steps)
median(steps_per_day$steps)


intervals <- group_by(df, interval)
activity <- summarize(intervals, steps = mean(steps, na.rm=TRUE))
with(activity, plot(interval, steps, type="l"))

head(arrange(activity,-steps),5)

missing<-is.na(df)
apply(missing,2, sum)

# use activity to replace missing values with mean values

get_mean_steps <- function(interval){
        avg <- activity[activity$interval==interval,2]
        avg
}


df2<-df

for(i in 1:dim(df2)[1]){
        if(is.na(df2[i,1])) df2[i,1] <- get_mean_steps(df2[i,3])
}



str(df2)str(d

days <- group_by(df2, date)
steps_per_day <-summarize( days, steps = sum(steps, na.rm=TRUE))
hist(steps_per_day$steps, breaks=30)

mean(steps_per_day$steps)
median(steps_per_day$steps)

df2$day <- weekdays(df2$datetime)
df2  <- mutate(df2, day = factor(1*(day=="Sunday" | day=="Saturday"),
                                     labels = c("weekday","weekend")))


intervals2 <- group_by(df2, interval)
activity2 <- summarize(intervals2, steps = mean(steps))

plot(activity$interval, activity$steps, type="l")
with(activity2, xyplot(  steps ~ interval, type="l"))


head(arrange(activity,-steps),5)
head(arrange(activity2,-steps),5)


```

d$corrected_steps <- d$steps
for (i in 1:dim(d)[1]){
  if(is.na(d[i,]$steps)){
    d[i,]$corrected_steps <- acs[acs$interval==d[i,]$interval,]$steps
  }
}

ints <- group_by(d, interval)
pp <- summarize(ints, steps = mean(steps, na.rm=TRUE), steps_corrected = mean(corrected_steps))
head(pp)
d$day <- weekdays(d$datetime)
head(d)
d <- mutate(d, day = factor(1*(day=="Saturday" | day=="Sunday"), labels=c("weekday","weekend")))
head(d)
day_interval <- group_by(d, day, interval)
head(day_interval)
spd <- summarize(day_interval, steps = mean(corrected_steps))


with(spd, xyplot(steps ~ interval | day, type="l",layout=c(1,2)))
